### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@58ea496e72ccfdf2bc9f32e816918cad1487d934 🚀
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@a0fd725e79a0798ad9322978f98274ab532f2319 🚀
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@081468149df4bbeb8db27e61cfa7a7be121e2af1 🚀

\texttt{NCBITaxonomy.jl} is a package designed to facilitate the
reconciliation and cleaning of taxonomic names, using a local copy of
the NCBI taxonomic backbone (Federhen 2012, Schoch et al. 2020); The
basic search functions are coupled with quality-of-life functions
including case-insensitive search and custom fuzzy string matching to
facilitate the amount of information that can be extracted automatically
while allowing efficient manual curation and inspection of results.
\texttt{NCBITaxonomy.jl} works with version 1.6 of the Julia programming
language (Bezanson et al. 2017), and relies on the Apache Arrow format
to store a local copy of the NCBI raw taxonomy files. The design of
\texttt{NCBITaxonomy.jl} has been inspired by similar efforts, like the
R package \texttt{taxadb} (Norman et al. 2020), which provides an
offline alternative to packages like \texttt{taxize} (Chamberlain and
Szöcs 2013).

\hypertarget{statement-of-need}{%
\section{Statement of need}\label{statement-of-need}}

Unambiguously identifying species is a far more challenging task than it
may appear. There are a vast number of reasons for this. Different
databases keep different taxonomic ``backbones,'' \emph{i.e.} different
data structures in which names are mapped to species, and organised in a
hierarchy. Not all names are unique identifiers to groups. For example,
\emph{Io} can either refer to a genus of plants from the aster family,
or to a genus of molluscs; the genus \emph{Mus} (of which the house
mouse \emph{Mus musculus} is a species), contains a sub-genus
\emph{also} named \emph{Mus}. Conversely, the same species can have
several names, which are valid synonyms: for example, the domestic cow
\emph{Bos taurus} admits \emph{Bos primigenius taurus} as a valid
synonym. Taxonomic nomenclature also changes regularly, with groups
being split, merged, or moved to a new position in the tree of life;
this is, notably, a common occurrence with viral taxonomy, each
subsequent version of which can differ markedly from the last; compare,
\emph{e.g} Lefkowitz et al. (2018) to Walker et al. (2020).

To add to the complexity, one must also consider that most taxa names
are at some point manually typed, which has the potential to introduce
additional mistakes in raw data; it is likely to expect that such
mistakes may arise when attempting to write down the (perfectly valid)
names of the bacterial isolate known as \emph{Myxococcus
llanfairpwllgwyngyllgogerychwyrndrobwllllantysiliogogogochensis}, or of
the crowned slaty flycatcher \emph{Griseotyrannus
aurantioatrocristatus}. These mistakes are more likely when dealing with
hyper-diverse samples, like plant census (Dauncey et al. 2016, Wagner
2016, Conti et al. 2021). In addition to binomial names, the same
species can be known by many vernacular (common) names, which are
language or even region-specific: \emph{Ovis aries}, for example, has
valid English vernaculars including lamb, sheep, wild sheep, and
domestic sheep.

@@

An up-to-date version of the documentation for \texttt{NCBITaxonomy.jl}
can be found online at
\url{https://docs.ecojulia.org/NCBITaxonomy.jl/stable/}, including
examples and a documentation of every method. The package is released
under the MIT license. Contributions can be made in the form of issues
(bug reports, questions, features suggestions) and pull requests. The
package can be checked out and installed anonymously from the central
Julia repository:

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J. et al. 2017. Julia: A Fresh Approach to Numerical
Computing. - SIAM Review 59: 65--98.

@@

\leavevmode\hypertarget{ref-Gibb2021DatPro}{}%
Gibb, R. et al. 2021. Data proliferation, reconciliation, and synthesis
in viral ecology. - bioRxiv: 2021.01.14.426572.

@@

>>>>

Unambiguously identifying species is a far more challenging task than it
may appear. There are a vast number of reasons for this. Different
databases keep different taxonomic ``backbones,'' \emph{i.e.} different
data structures in which names are mapped to species, and organised in a
hierarchy. Not all names are unique identifiers to groups. For example,
\emph{Io} can either refer to a genus of plants from the aster family,
or to a genus of molluscs; the genus \emph{Mus} (of which the house
mouse \emph{Mus musculus} is a species), contains a sub-genus
\emph{also} named \emph{Mus} (within which \emph{Mus musculus} is
located). Conversely, the same species can have several names, which are
valid synonyms: for example, the domestic cow \emph{Bos taurus} admits
\emph{Bos primigenius taurus} as a valid synonym. In addition to
binomial names, the same species can be known by many vernacular
(common) names, which are language or even region-specific: \emph{Ovis
aries}, for example, has valid English vernaculars including lamb,
sheep, wild sheep, and domestic sheep.

Finally, taxonomic nomenclature changes regularly, with groups being
split, merged, or moved to a new position in the tree of life; this is,
notably, a common occurrence with viral taxonomy, each subsequent
version of which can differ markedly from the last; compare, \emph{e.g}
Lefkowitz et al. (2018) to Walker et al. (2020), where entire viral
sub-trees were split, re-organized, and created within just two years.
These taxonomic changes have profound implications for the way we
perceive biodiversity at global scales (Meier 2009), to the point were
taxonomic revisions should sometimes be actively conducted to improve
\emph{e.g.} conservation outcomes (Melville et al. 2021).

To add to the complexity, one must also consider that most taxa names
are at some point manually typed, which has the potential to introduce
additional mistakes in raw data; it is likely to expect that such
mistakes may arise when attempting to write down the (perfectly valid)
names of the bacterial isolate known as \emph{Myxococcus
llanfairpwllgwyngyllgogerychwyrndrobwllllantysiliogogogochensis}, or of
the crowned slaty flycatcher \emph{Griseotyrannus
aurantioatrocristatus}. These mistakes are more likely when dealing with
hyper-diverse samples (demanding to memorize more names), like plant
census (Dauncey et al. 2016, Wagner 2016, Conti et al. 2021); when
dealing with multiple investigators with different knowledge of the
taxonomy; and as a result of the estimated error in any data entry
exercice, which other fields estimate at up to about 5\% (Barchard and
Pace 2011).

@@

An up-to-date version of the documentation for \texttt{NCBITaxonomy.jl}
can be found online from the EcoJulia documentation hub
\href{https://docs.ecojulia.org/NCBITaxonomy.jl/stable/}{https://docs.ecojulia.org/},
including examples and in-line documentation of every method. The
package is released under the MIT license. Contributions can be made in
the form of issues (bug reports, questions, features suggestions) and
pull requests. The package can be checked out and installed anonymously
from the central Julia repository:

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Barchard2011PreHum}{}%
Barchard, K. and Pace, L. 2011. Preventing human error: The impact of
data entry methods on data accuracy and statistical results. - Computers
in Human Behavior 27: 1834--1839.

\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J. et al. 2017. Julia: A Fresh Approach to Numerical
Computing. - SIAM Review 59: 65--98.

@@

\leavevmode\hypertarget{ref-Gibb2021DatPro}{}%
Gibb, R. et al. 2021. Data Proliferation, Reconciliation, and Synthesis
in Viral Ecology. - BioScience in press.

@@

\leavevmode\hypertarget{ref-Meier2009BioRes}{}%
Meier, R. 2009. Biodiversity Research Based on Taxonomic Revisions - A
Tale of Unrealized Opportunities -. - In: pp. 323--345.

\leavevmode\hypertarget{ref-Melville2021RetApp}{}%
Melville, J. et al. 2021. A return-on-investment approach for
prioritization of rigorous taxonomic research needed to inform responses
to the biodiversity crisis. - PLOS Biology 19: e3001210.

### 2022-06-27 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@58ea496e72ccfdf2bc9f32e816918cad1487d934 🚀
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@a0fd725e79a0798ad9322978f98274ab532f2319 🚀
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@081468149df4bbeb8db27e61cfa7a7be121e2af1 🚀

\hypertarget{statement-of-need}{%
\section{Statement of need}\label{statement-of-need}}

Unambiguously identifying species is a far more challenging task than it
may appear. There are a vast number of reasons for this. Different
databases keep different taxonomic ``backbones,'' \emph{i.e.} different
data structures in which names are mapped to species, and organised in a
hierarchy. Not all names are unique identifiers to groups. For example,
\emph{Io} can either refer to a genus of plants from the aster family,
or to a genus of molluscs; the genus \emph{Mus} (of which the house
mouse \emph{Mus musculus} is a species), contains a sub-genus
\emph{also} named \emph{Mus}. Conversely, the same species can have
several names, which are valid synonyms: for example, the domestic cow
\emph{Bos taurus} admits \emph{Bos primigenius taurus} as a valid
synonym. Taxonomic nomenclature also changes regularly, with groups
being split, merged, or moved to a new position in the tree of life;
this is, notably, a common occurrence with viral taxonomy, each
subsequent version of which can differ markedly from the last; compare,
\emph{e.g} Lefkowitz et al. (2018) to Walker et al. (2020).

To add to the complexity, one must also consider that most taxa names
are at some point manually typed, which has the potential to introduce
additional mistakes in raw data; it is likely to expect that such
mistakes may arise when attempting to write down the (perfectly valid)
names of the bacterial isolate known as \emph{Myxococcus
llanfairpwllgwyngyllgogerychwyrndrobwllllantysiliogogogochensis}, or of
the crowned slaty flycatcher \emph{Griseotyrannus
aurantioatrocristatus}. These mistakes are more likely when dealing with
hyper-diverse samples, like plant census (Dauncey et al. 2016, Wagner
2016, Conti et al. 2021). In addition to binomial names, the same
species can be known by many vernacular (common) names, which are
language or even region-specific: \emph{Ovis aries}, for example, has
valid English vernaculars including lamb, sheep, wild sheep, and
domestic sheep.

>>>>

Unambiguously identifying species is a far more challenging task than it
may appear. There are a vast number of reasons for this. Different
databases keep different taxonomic ``backbones,'' \emph{i.e.} different
data structures in which names are mapped to species, and organised in a
hierarchy. Not all names are unique identifiers to groups. For example,
\emph{Io} can either refer to a genus of plants from the aster family,
or to a genus of molluscs; the genus \emph{Mus} (of which the house
mouse \emph{Mus musculus} is a species), contains a sub-genus
\emph{also} named \emph{Mus} (within which \emph{Mus musculus} is
located). Conversely, the same species can have several names, which are
valid synonyms: for example, the domestic cow \emph{Bos taurus} admits
\emph{Bos primigenius taurus} as a valid synonym. In addition to
binomial names, the same species can be known by many vernacular
(common) names, which are language or even region-specific: \emph{Ovis
aries}, for example, has valid English vernaculars including lamb,
sheep, wild sheep, and domestic sheep.

Finally, taxonomic nomenclature changes regularly, with groups being
split, merged, or moved to a new position in the tree of life; this is,
notably, a common occurrence with viral taxonomy, each subsequent
version of which can differ markedly from the last; compare, \emph{e.g}
Lefkowitz et al. (2018) to Walker et al. (2020), where entire viral
sub-trees were split, re-organized, and created within just two years.
These taxonomic changes have profound implications for the way we
perceive biodiversity at global scales (Meier 2009), to the point were
taxonomic revisions should sometimes be actively conducted to improve
\emph{e.g.} conservation outcomes (Melville et al. 2021).

To add to the complexity, one must also consider that most taxa names
are at some point manually typed, which has the potential to introduce
additional mistakes in raw data; it is likely to expect that such
mistakes may arise when attempting to write down the (perfectly valid)
names of the bacterial isolate known as \emph{Myxococcus
llanfairpwllgwyngyllgogerychwyrndrobwllllantysiliogogogochensis}, or of
the crowned slaty flycatcher \emph{Griseotyrannus
aurantioatrocristatus}. These mistakes are more likely when dealing with
hyper-diverse samples (demanding to memorize more names), like plant
census (Dauncey et al. 2016, Wagner 2016, Conti et al. 2021); when
dealing with multiple investigators with different knowledge of the
taxonomy; and as a result of the estimated error in any data entry
exercice, which other fields estimate at up to about 5\% (Barchard and
Pace 2011).
